---
title: "Metabolism Decision Tree"
author: "Alice Carter"
date: "8/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include=F)
```
## Modeling Metabolism in New Hope Creek

#### What decisions do we make when modeling metabolism?
#### How do these influence our results? 

To explore this, we will use example data from New Hope Creek, NC 

  

```{r}
library(tidyverse)
library(dygraphs)
library(xts)
library(lubridate)



setwd(metab_projdir)
dat<- readRDS("data/metabolism/condensed/condensed_NHC.rds")
nreg <- read_csv("data/siteData/nightreg_NHC.csv")
dat<- dat$data
DO <- select(dat, solar.time, DO.obs, DO.sat, discharge)
DO$logQ <- log(DO$discharge)
DO$DO_pctsat <- DO$DO.obs/DO$DO.sat

DO_xts <- xts(DO[,5:6], order.by = DO$solar.time)
```

```{r DO_xts, include=T}
dygraph(DO_xts, main = "New Hope Creek")%>%
  dyRangeSelector()

```





I visually selected three days of data that capture the range of dynamics observed in NHC, including:

  1. DO swings high, near saturation: 2020 Mar 7-9

  2. DO swings low, near saturation:  2020 Feb 8-10

  3. DO swings high, undersaturated;  2019 Sept 21-23

  4. DO swings low, undersaturated:   2019 Sept 28-30


```{r dat, include=T}
start_date <- c("2020-03-07", "2020-02-08", "2019-09-21", "2019-09-28")
start_datetime <- ymd_hms(paste(start_date, "04:00:00"))
end_datetime <- start_datetime+3*24*60*60
dat <- select(dat, -DO.mod)
dat <- left_join(dat, nreg[,1:2])

dat_1 <- dplyr::filter(dat, solar.time>=start_datetime[1] & solar.time <= end_datetime[1])
dat_2 <- dplyr::filter(dat, solar.time>=start_datetime[2] & solar.time <= end_datetime[2])
dat_3 <- dplyr::filter(dat, solar.time>=start_datetime[3] & solar.time <= end_datetime[3])
dat_4 <- dplyr::filter(dat, solar.time>=start_datetime[4] & solar.time <= end_datetime[4])

par(mfrow = c(7,2), mar=c(0,4,0,4), oma = c(4,4,4,4))
plot(dat_1$solar.time, dat_1$temp.water, type="l", ylim = c(5,25), lwd=2, ylab = "temp", xaxt='n')
par(new=T)
plot(dat_1$solar.time, dat_1$light, type="l", ylim = c(0,1800), col = "orange", lwd=2, axes=F, ylab="", xlab="")
axis(4)
mtext("light", 4, 2.2, cex=.7)
mtext("1: High DO swings, near saturation", 3, line=1.5, cex=.9)

plot(dat_2$solar.time, dat_2$temp.water, type="l", ylim = c(5,25), lwd=2, ylab = "temp", xaxt='n')
par(new=T)
plot(dat_2$solar.time, dat_2$light, type="l", ylim = c(0,1800), col = "orange", lwd=2, axes=F, ylab="", xlab="")
axis(4)
mtext("light", 4, 2.2, cex=.7)
mtext("2: Low DO swings, near saturation", 3, line=1.5, cex=.9)

plot(dat_1$solar.time, dat_1$DO.obs/dat_1$DO.sat, type="l", col="darkblue", lwd=2, ylab="DO %sat", ylim = c(.9,1.2), xaxt="n")

plot(dat_2$solar.time, dat_2$DO.obs/dat_2$DO.sat, type="l", col="darkblue", lwd=2, ylab="DO %sat", ylim = c(.9,1.2), xaxt="n")

par(mar = c(2,4,0,4))
plot(dat_1$solar.time, dat_1$discharge, type="l", log="y", lwd=2,col="darkgrey", ylab="Q", ylim = c(0.001, 500))

plot(dat_2$solar.time, dat_2$discharge, type="l", log="y", lwd=2,col="darkgrey", ylab="Q", ylim = c(0.001, 500))

plot(1,1, type="n", axes=F, xaxt='n', yaxt='n', xlab="", ylab="")
plot(1,1, type="n", axes=F, xaxt='n', yaxt='n', xlab="", ylab="")

par(mar = c(0,4,0,4))
plot(dat_3$solar.time, dat_3$temp.water, type="l", ylim = c(5,25), lwd=2, ylab = "temp", xaxt='n')
par(new=T)
plot(dat_3$solar.time, dat_3$light, type="l", ylim = c(0,1800), col = "orange", lwd=2, axes=F, ylab="", xlab="")
axis(4)
mtext("light", 4, 2.2, cex=.7)
mtext("3: High DO swings, undersaturated", 3, line=1.5, cex=.9)

plot(dat_4$solar.time, dat_4$temp.water, type="l", ylim = c(5,25), lwd=2, ylab = "temp", xaxt='n')
par(new=T)
plot(dat_4$solar.time, dat_4$light, type="l", ylim = c(0,1800), col = "orange", lwd=2, axes=F, ylab="", xlab="")
axis(4)
mtext("light", 4, 2.2, cex=.7)
mtext("4: Low DO swings, undersaturated", 3, line=1.5, cex=.9)

plot(dat_3$solar.time, dat_3$DO.obs/dat_3$DO.sat, type="l", col="darkblue", lwd=2, ylab="DO %sat", ylim = c(0.08,0.38), xaxt="n")

plot(dat_4$solar.time, dat_4$DO.obs/dat_4$DO.sat, type="l", col="darkblue", lwd=2, ylab="DO %sat", ylim = c(0,0.3), xaxt="n")

par(mar = c(2,4,0,4))
plot(dat_3$solar.time, dat_3$discharge, type="l", log="y", lwd=2,col="darkgrey", ylab="Q", ylim = c(0.001, 500))

plot(dat_4$solar.time, dat_4$discharge, type="l", log="y", lwd=2,col="darkgrey", ylab="Q", ylim = c(0.001, 500))

```




### Running Stream Metabolizer

Decision 1: Bayesian model vs Maximum Likelihood

Decision 2: 

  a) pooled K based on Q

  b) fixed K based on night regression



Running a basic MLE or Bayes model does not require discharge, only depth*

  * this is actually a decision I bypassed. How do we calculate depth in streams? I am currently using an empirical equation built into stream metabolizer

```{r}
library(streamMetabolizer)

dates <- unique(dat_1$date)
mle_name <- mm_name(type="mle")
mle_specs <- specs(mle_name)
tmp <- select(dat_1, -discharge, -date, -K600)
mle_mm<- metab(mle_specs, tmp)
mle_preds <- mle_mm@fit


bayes_np_name <- mm_name(type="bayes", pool_K600="none")
bayes_np_specs <- specs(bayes_np_name)
bayes_np_mm<- metab(bayes_np_specs, tmp)
bnp_preds <- bayes_np_mm@fit$daily[,c(1,7,5,9,17,15,19,57,55,59)]
colnames(bnp_preds) <- c("date", "GPP","GPP.lower","GPP.upper", "ER","ER.lower","ER.upper","K600","K600.lower","K600.upper")

bayes_name <- mm_name(type="bayes", pool_K600="binned", err_obs_iid = TRUE,
                      err_proc_iid = TRUE, err_proc_acor = FALSE)
bayes_specs <- specs(bayes_name)
tmp <- tibble(select(dat_1, -date, -K600))
bayes_mm <- metab(bayes_specs, tmp)
bp_preds <- bayes_mm@fit$daily[,c(1,7,5,9,17,15,19,57,55,59)]
colnames(bp_preds) <- c("date", "GPP","GPP.lower","GPP.upper", "ER","ER.lower","ER.upper","K600","K600.lower","K600.upper")

 
KQ <- filter(nreg, date %in% dates)
# bayes_name<- mm_name(type="bayes", pool_K600="binned_sdzero", err_obs_iid = TRUE,
#                       err_proc_iid = TRUE, err_proc_acor = FALSE)
# bayes_specs <- specs(bayes_name)
# bayes_specs$K600_lnQ_nodes_centers <- log(KQ$discharge_m3s)
# bayes_specs$K600_lnQ_nodes_meanlog <- log(KQ$K600)
# bayes_specs$K600_lnQ_nodes_sdlog <- rep(.1,3)
# bayes_K_prior_mm <- metab(bayes_specs, tmp)
# bkpr_preds <- bayes_K_prior_mm@fit$daily[,c(1,7,5,9,17,15,19,57,55,59)]
# colnames(bkpr_preds) <- c("date", "GPP","GPP.lower","GPP.upper", "ER","ER.lower","ER.upper","K600","K600.lower","K600.upper")


```


```{r, include=T}
par(mfrow=c(3, 1), mar = c(0,4,0,2), oma = c(5,1,3,1))
plot(mle_preds$date, mle_preds$GPP.daily, type="l", lwd=2, ylab="GPP", xaxt="n", ylim = c(1.06,1.29))
mtext("Scenario 1: High DO swings, near saturation", 3,1.5)
lines(bnp_preds$date, bnp_preds$GPP,  lwd=2, col="forestgreen")
lines(bp_preds$date, bp_preds$GPP, lwd=2, lty=2,col="forestgreen")
#lines(bkpr_preds$date, bkpr_preds$GPP, lwd=2, lty=3,col="forestgreen",)
legend("topright", c("mle","bayes","bayes K pooled","night reg K" ), col=c("black","grey","grey","black"),
       lty=c(1,1,2,NA), lwd=2, pch=c(NA,NA,NA,1), bty="n",ncol=2)

plot(mle_preds$date, mle_preds$ER.daily, type="l", lwd=2, ylab="ER", xaxt="n", ylim = c(-0.35,-.1))
lines(bnp_preds$date, bnp_preds$ER, lwd=2, col="sienna")
lines(bp_preds$date, bp_preds$ER,  lwd=2, lty=2,col="sienna")
#lines(bkpr_preds$date, bkpr_preds$ER, lwd=2, lty=3,col="sienna")

plot(mle_preds$date, mle_preds$K600.daily, type="l", lwd=2, ylab="K600", ylim = c(7.2,9.4))
lines(bnp_preds$date, bnp_preds$K600, lwd=2, col="steelblue")
lines(bp_preds$date, bp_preds$K600, lwd=2, lty=2,col="steelblue")
#lines(bkpr_preds$date, bkpr_preds$K600, lwd=2, lty=3,col="steelblue")
points(KQ$date, KQ$K600, pch=1)




```



```{r}
dates <- unique(dat_2$date)
mle_name <- mm_name(type="mle")
mle_specs <- specs(mle_name)
tmp <- select(dat_2, -discharge, -date, -K600)
mle_mm<- metab(mle_specs, tmp)
mle_preds <- mle_mm@fit


bayes_np_name <- mm_name(type="bayes", pool_K600="none")
bayes_np_specs <- specs(bayes_np_name)
bayes_np_mm<- metab(bayes_np_specs, tmp)
bnp_preds <- bayes_np_mm@fit$daily[,c(1,7,5,9,17,15,19,57,55,59)]
colnames(bnp_preds) <- c("date", "GPP","GPP.lower","GPP.upper", "ER","ER.lower","ER.upper","K600","K600.lower","K600.upper")

bayes_name <- mm_name(type="bayes", pool_K600="binned", err_obs_iid = TRUE,
                      err_proc_iid = TRUE, err_proc_acor = FALSE)
bayes_specs <- specs(bayes_name)
tmp <- tibble(select(dat_2, -date, -K600))
bayes_mm <- metab(bayes_specs, tmp)
bp_preds <- bayes_mm@fit$daily[,c(1,7,5,9,17,15,19,57,55,59)]
colnames(bp_preds) <- c("date", "GPP","GPP.lower","GPP.upper", "ER","ER.lower","ER.upper","K600","K600.lower","K600.upper")

 
KQ <- filter(nreg, date %in% dates)
bayes_name<- mm_name(type="bayes", pool_K600="binned_sdzero", err_obs_iid = TRUE,
                      err_proc_iid = TRUE, err_proc_acor = FALSE)
bayes_specs <- specs(bayes_name)
bayes_specs$K600_lnQ_nodes_centers <- log(KQ$discharge_m3s)
bayes_specs$K600_lnQ_nodes_meanlog <- log(KQ$K600)
bayes_specs$K600_lnQ_nodes_sdlog <- rep(.1,3)
bayes_K_prior_mm <- metab(bayes_specs, tmp)
bkpr_preds <- bayes_K_prior_mm@fit$daily[,c(1,7,5,9,17,15,19,57,55,59)]
colnames(bkpr_preds) <- c("date", "GPP","GPP.lower","GPP.upper", "ER","ER.lower","ER.upper","K600","K600.lower","K600.upper")


```


```{r, include=T}
par(mfrow=c(3, 1), mar = c(0,4,0,2), oma = c(5,1,3,1))
plot(mle_preds$date, mle_preds$GPP.daily, type="l", lwd=2, ylab="GPP", xaxt="n")
mtext("Scenario 2: Low DO swings, near saturation", 3,1.5)
lines(bnp_preds$date, bnp_preds$GPP,  lwd=2, col="forestgreen")
lines(bp_preds$date, bp_preds$GPP, lwd=2, lty=2,col="forestgreen")
lines(bkpr_preds$date, bkpr_preds$GPP, lwd=2, lty=3,col="forestgreen",)

legend("topright", c("mle","bayes","bayes K pooled","bayes k fixed","night reg K"  ), col=c("black","grey","grey","grey","black"),
       lty=c(1,1,2,3,NA), lwd=2, pch=c(NA,NA,NA,NA,1), bty="n",ncol=2)

plot(mle_preds$date, mle_preds$ER.daily, type="l", lwd=2, ylab="ER", xaxt="n")
lines(bnp_preds$date, bnp_preds$ER, lwd=2, col="sienna")
lines(bp_preds$date, bp_preds$ER,  lwd=2, lty=2,col="sienna")
lines(bkpr_preds$date, bkpr_preds$ER, lwd=2, lty=3,col="sienna")

plot(mle_preds$date, mle_preds$K600.daily, type="l", lwd=2, ylab="K600", ylim = c(8,50))
lines(bnp_preds$date, bnp_preds$K600, lwd=2, col="steelblue")
lines(bp_preds$date, bp_preds$K600, lwd=2, lty=2,col="steelblue")
lines(bkpr_preds$date, bkpr_preds$K600, lwd=2, lty=3,col="steelblue")
points(KQ$date, KQ$K600, pch=1)


```




```{r}
dates <- unique(dat_3$date)
mle_name <- mm_name(type="mle")
mle_specs <- specs(mle_name)
tmp <- select(dat_3, -discharge, -date, -K600)
mle_mm<- metab(mle_specs, tmp)
mle_preds <- mle_mm@fit


bayes_np_name <- mm_name(type="bayes", pool_K600="none")
bayes_np_specs <- specs(bayes_np_name)
bayes_np_mm<- metab(bayes_np_specs, tmp)
bnp_preds <- bayes_np_mm@fit$daily[,c(1,7,5,9,17,15,19,57,55,59)]
colnames(bnp_preds) <- c("date", "GPP","GPP.lower","GPP.upper", "ER","ER.lower","ER.upper","K600","K600.lower","K600.upper")

bayes_name <- mm_name(type="bayes", pool_K600="binned", err_obs_iid = TRUE,
                      err_proc_iid = TRUE, err_proc_acor = FALSE)
bayes_specs <- specs(bayes_name)
tmp <- tibble(select(dat_3, -date, -K600))
bayes_mm <- metab(bayes_specs, tmp)
bp_preds <- bayes_mm@fit$daily[,c(1,7,5,9,17,15,19,57,55,59)]
colnames(bp_preds) <- c("date", "GPP","GPP.lower","GPP.upper", "ER","ER.lower","ER.upper","K600","K600.lower","K600.upper")

 
KQ <- filter(nreg, date %in% dates)
# bayes_name<- mm_name(type="bayes", pool_K600="binned_sdzero", err_obs_iid = TRUE,
#                       err_proc_iid = TRUE, err_proc_acor = FALSE)
# bayes_specs <- specs(bayes_name)
# bayes_specs$K600_lnQ_nodes_centers <- log(KQ$discharge_m3s)
# bayes_specs$K600_lnQ_nodes_meanlog <- log(KQ$K600)
# bayes_specs$K600_lnQ_nodes_sdlog <- rep(.1,3)
# bayes_K_prior_mm <- metab(bayes_specs, tmp)
# bkpr_preds <- bayes_K_prior_mm@fit$daily[,c(1,7,5,9,17,15,19,57,55,59)]
# colnames(bkpr_preds) <- c("date", "GPP","GPP.lower","GPP.upper", "ER","ER.lower","ER.upper","K600","K600.lower","K600.upper")


```


```{r, include=T}
par(mfrow=c(3, 1), mar = c(0,4,0,2), oma = c(5,1,3,1))
plot(mle_preds$date, mle_preds$GPP.daily, type="l", lwd=2, ylab="GPP", xaxt="n")
mtext("Scenario 3: High DO swings, undersaturated", 3,1.5)
lines(bnp_preds$date, bnp_preds$GPP,  lwd=2, col="forestgreen")
lines(bp_preds$date, bp_preds$GPP, lwd=2, lty=2,col="forestgreen")
#lines(bkpr_preds$date, bkpr_preds$GPP, lwd=2, lty=3,col="forestgreen",)

legend("topleft", c("mle","bayes","bayes K pooled","night reg K" ), col=c("black","grey","grey","black"),
       lty=c(1,1,2,NA), lwd=2, pch=c(NA,NA,NA,1), bty="n",ncol=2)

plot(mle_preds$date, mle_preds$ER.daily, type="l", lwd=2, ylab="ER", xaxt="n")
lines(bnp_preds$date, bnp_preds$ER, lwd=2, col="sienna")
lines(bp_preds$date, bp_preds$ER,  lwd=2, lty=2,col="sienna")
#lines(bkpr_preds$date, bkpr_preds$ER, lwd=2, lty=3,col="sienna")

plot(mle_preds$date, mle_preds$K600.daily, type="l", lwd=2, ylab="K600")#, ylim = c(8,50))
lines(bnp_preds$date, bnp_preds$K600, lwd=2, col="steelblue")
lines(bp_preds$date, bp_preds$K600, lwd=2, lty=2,col="steelblue")
#lines(bkpr_preds$date, bkpr_preds$K600, lwd=2, lty=3,col="steelblue")
points(KQ$date, KQ$K600, pch=1)


```






```{r}

dates <- unique(dat_4$date)
mle_name <- mm_name(type="mle")
mle_specs <- specs(mle_name)
tmp <- select(dat_4, -discharge, -date, -K600)
mle_mm<- metab(mle_specs, tmp)
mle_preds <- mle_mm@fit


bayes_np_name <- mm_name(type="bayes", pool_K600="none")
bayes_np_specs <- specs(bayes_np_name)
bayes_np_mm<- metab(bayes_np_specs, tmp)
bnp_preds <- bayes_np_mm@fit$daily[,c(1,7,5,9,17,15,19,57,55,59)]
colnames(bnp_preds) <- c("date", "GPP","GPP.lower","GPP.upper", "ER","ER.lower","ER.upper","K600","K600.lower","K600.upper")

bayes_name <- mm_name(type="bayes", pool_K600="binned", err_obs_iid = TRUE,
                      err_proc_iid = TRUE, err_proc_acor = FALSE)
bayes_specs <- specs(bayes_name)
tmp <- tibble(select(dat_4, -date, -K600))
bayes_mm <- metab(bayes_specs, tmp)
bp_preds <- bayes_mm@fit$daily[,c(1,7,5,9,17,15,19,57,55,59)]
colnames(bp_preds) <- c("date", "GPP","GPP.lower","GPP.upper", "ER","ER.lower","ER.upper","K600","K600.lower","K600.upper")

 
KQ <- filter(nreg, date %in% dates)
# bayes_name<- mm_name(type="bayes", pool_K600="binned_sdzero", err_obs_iid = TRUE,
#                       err_proc_iid = TRUE, err_proc_acor = FALSE)
# bayes_specs <- specs(bayes_name)
# bayes_specs$K600_lnQ_nodes_centers <- log(KQ$discharge_m3s)
# bayes_specs$K600_lnQ_nodes_meanlog <- log(KQ$K600)
# bayes_specs$K600_lnQ_nodes_sdlog <- rep(.1,3)
# bayes_K_prior_mm <- metab(bayes_specs, tmp)
# bkpr_preds <- bayes_K_prior_mm@fit$daily[,c(1,7,5,9,17,15,19,57,55,59)]
# colnames(bkpr_preds) <- c("date", "GPP","GPP.lower","GPP.upper", "ER","ER.lower","ER.upper","K600","K600.lower","K600.upper")


```


```{r, include=T}
par(mfrow=c(3, 1), mar = c(0,4,0,2), oma = c(5,1,3,1))
plot(mle_preds$date, mle_preds$GPP.daily, type="l", lwd=2, ylab="GPP", xaxt="n", ylim = c(-0.02,.04))
abline(h=0, lwd=.5)
mtext("Scenario 4: low DO swings, undersaturated", 3,1.5)
lines(bnp_preds$date, bnp_preds$GPP,  lwd=2, col="forestgreen")
lines(bp_preds$date, bp_preds$GPP, lwd=2, lty=2,col="forestgreen")
#lines(bkpr_preds$date, bkpr_preds$GPP, lwd=2, lty=3,col="forestgreen",)

legend("topleft", c("mle","bayes","bayes K pooled","night reg K" ), col=c("black","grey","grey","black"),
       lty=c(1,1,2,NA), lwd=2, pch=c(NA,NA,NA,1), bty="n",ncol=2)

plot(mle_preds$date, mle_preds$ER.daily, type="l", lwd=2, ylab="ER", xaxt="n")
lines(bnp_preds$date, bnp_preds$ER, lwd=2, col="sienna")
lines(bp_preds$date, bp_preds$ER,  lwd=2, lty=2,col="sienna")
#lines(bkpr_preds$date, bkpr_preds$ER, lwd=2, lty=3,col="sienna")

plot(mle_preds$date, mle_preds$K600.daily, type="l", lwd=2, ylab="K600")#, ylim = c(8,50))
lines(bnp_preds$date, bnp_preds$K600, lwd=2, col="steelblue")
lines(bp_preds$date, bp_preds$K600, lwd=2, lty=2,col="steelblue")
#lines(bkpr_preds$date, bkpr_preds$K600, lwd=2, lty=3,col="steelblue")
points(KQ$date, KQ$K600, pch=1)

```

